---
title: "ECCA Sampling"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(dplyr)
library(units)
library(lwgeom)
library(stringr)
library(ggplot2)
library(maptiles)

CRS_FEET = 2927
CRS_LATLONG = 4326
```

# Segment Generation

## Functions

```{r}
add_mileposts_to_segment <- function (data) {
  data_segments <- st_cast(data, 'LINESTRING')
  lengths <- as.numeric(st_length(st_transform(data_segments, CRS_FEET)) / 5280)
  mp.starts <- c(0, head(lengths, -1))
  mp.ends <- mp.starts + lengths
  data_segments |> mutate(BARM = mp.starts, EARM = mp.ends)
}

extract_milepost_segments <- function (data, start_mp, end_mp) {
  filtered_data <- data |> filter(BARM < {{ end_mp }} & EARM > {{ start_mp }})
  
  clipped_list <- lapply(1:nrow(filtered_data), function(i) {
    chunk_by_mileposts(filtered_data[i, ], start_mp, end_mp)
  })
  
  area <- st_as_sf(data.frame(geometry = do.call(c, clipped_list)), crs = st_crs(data)) |>
    st_transform(CRS_FEET)
  
  if (nrow(area) > 1) {
    area <- st_line_merge(st_union(area))
  }
  area
}

chunk_by_mileposts <- function (row, start_mp, end_mp) {
  row_start <- row$BARM # beginning of accumulated range of mileage
  row_end <- row$EARM # ending of accumulated range of mileage
  row_length <- row_end - row_start
  
  if (row_start >= end_mp || row_end < start_mp) {
    return (NULL)
  }
  
  new_start <- max(row_start, start_mp)
  new_end <- min(row_end, end_mp)
  
  if (new_start >= new_end) {
    return (NULL)
  }
  
  t_start <- (new_start - row_start) / row_length
  t_end <- (new_end - row_start) / row_length
  st_linesubstring(row$geometry, t_start, t_end)
}

create_segments <- function (area, segment_length=10^10) {
  line <- area[1,]
  line_length <- as.numeric(st_length(line))

  fractions <- fixed_length_fractions(line_length, segment_length)
  n_segments <- nrow(fractions)
  
  lapply(seq_len(n_segments), function(i) {
    st_linesubstring(line, from = fractions$starts[i], to = fractions$ends[i])
  }) |>
    data.frame() |>
    t() |>
    st_sfc() |>
    st_as_sf(crs = st_crs(line))
}

# this forces all segments the same length and distributes the remaining gap as equal-sized chunks between all the real segments
fixed_length_fractions <- function (line_length, segment_length) {
  total_gap <- line_length %% segment_length
  n_segments <- floor(line_length / segment_length)
  gap_step <- total_gap / (n_segments - 1)
  step_size = segment_length + gap_step
  
  segment_starts <- seq(0, line_length, by=step_size)
  segment_ends <- pmin(segment_starts + segment_length, line_length)
  
  fraction_starts <- segment_starts / line_length
  fraction_ends <- segment_ends / line_length
  data.frame(starts=fraction_starts, ends=fraction_ends)
}

make_data_frame <- function (segments, start_mp, end_mp, area_label, segment_prefix=NULL) {
  segments$area <- area_label
  segments$area.id <- str_c(segment_prefix, seq_len(nrow(segments)), sep='-')
  
  mp_seq <- seq(start_mp, end_mp, length.out = nrow(segments) + 1)
  segments$start.milepost <- head(mp_seq, -1)
  segments$end.milepost   <- tail(mp_seq, -1)
  
  segments <- st_transform(segments, CRS_LATLONG)
  start_points <- st_startpoint(segments) |> st_coordinates()
  segments$start.latitude <- as.character(start_points[,1])
  segments$start.longitude <- as.character(start_points[,2])
  segments$start.maplink <- paste0("https://www.google.com/maps?q=", segments$start.lon, ",", segments$start.lat)
  
  end_points <- st_endpoint(segments) |> st_coordinates()
  segments$end.latitude <- as.character(end_points[,1])
  segments$end.longitude <- as.character(end_points[,2])
  segments$end.maplink <- paste0("https://www.google.com/maps?q=", segments$end.lon, ",", segments$end.lat)

  segments
}
```

## Highway Stretches (Increasing/Decreasing)

The LRS map includes separate segments for increasing mileposts (south side, eastbound) and decreasing mileposts (north side, westbound).

```{r}
make_highway_segments <- function (data, start_mp, end_mp, segment_size, label, area_prefix) {
  extract_milepost_segments(data, start_mp, end_mp) |>
    create_segments(segment_size) |>
    make_data_frame(start_mp, end_mp, label, area_prefix)
}

highway_segment_size = 600 # feet
highway_file <- './data/WSDOT_-_State_Route_Linear_Referencing_System_(LRS)_Current.shp'
routes <- st_read(highway_file)

wa14_mp_inc <- routes |>
  filter(RouteIdent == '014i') |> # Eastbound/south side (mileposts increasing)
  add_mileposts_to_segment()

wa14_mp_dec <- routes |>
  filter(RouteIdent == '014d') |> # Westbound/north side (mileposts decreasing)
  add_mileposts_to_segment()

# Mile post boundaries are based on ECCA agreements with the state
area1N <- make_highway_segments(wa14_mp_dec, 10.0, 11.99, highway_segment_size, 'Highway: Camas-N', 'H1')
area1S <- make_highway_segments(wa14_mp_inc, 10.09, 10.58, highway_segment_size, 'Highway: Camas-S', 'H2')
area2S <- make_highway_segments(wa14_mp_inc, 13.75, 14.35, highway_segment_size, 'Highway: Washougal-SW-Extra', 'H3')
area3N <- make_highway_segments(wa14_mp_dec, 15.05, 17.0, highway_segment_size, 'Highway: Washougal-N', 'H4')
area3S <- make_highway_segments(wa14_mp_inc, 15.05, 17.0, highway_segment_size, 'Highway: Washgoual-S', 'H5')
```

## Ramps

```{r}
make_ramp_segments <- function (data, id, start_mp, end_mp, label) {
  area <- data |> filter(RouteID == id) |>
    add_mileposts_to_segment() |>
    extract_milepost_segments(start_mp, end_mp) |>
    create_segments() |>
    make_data_frame(start_mp, end_mp, label) |>
    mutate(area.id = id)
}

is_on_ramp <- function (ramp) {
  ramp_code <- (ramp$area.id |> strsplit(''))[[1]][4]
  ramp_code == 'Q' | ramp_code == 'S'
}

is_off_ramp <- function (ramp) {
  !is_on_ramp(ramp)
}

ramp_file <- './data/WSDOT_-_State_Route_Ramp_(1%3A24K)_Current.shp'
ramps <- st_read(ramp_file)


vancouver_ramps <- rbind(
  make_ramp_segments(ramps, '014Q101046', 0.01, 0.21, 'Ramp: Vancouver S on-ramp'),
  make_ramp_segments(ramps, '014R101037', 0.06, 0.27, 'Ramp: Vancouver N off-ramp')
)

camas_ramps <- rbind(
  make_ramp_segments(ramps, '014S101194', 0.13, 0.45, 'Ramp: Camas N on-ramp'),
  make_ramp_segments(ramps, '014P101203', 0, 0.34, 'Ramp: Camas S off-ramp')
)

washgoual_on_ramp <- make_ramp_segments(ramps, '014Q101544', 0, 0.26, 'Ramp: Washougal S on-ramp') # eastern pair
washougal_off_ramp <- make_ramp_segments(ramps, '014P101430', 0.07, 0.38, 'Ramp: Washougal S off-ramp') # western pair, part of "unofficial" stretch
washougal_ramps <- rbind(washgoual_on_ramp, washougal_off_ramp)
```

# Sampling

The official sampling has already been run so this shouldn't be needed other than for reference.

```{r}
# These are the officially sampled areas
sample_area_ids = c('H1-8', 'H2-1', 'H3-3', 'H4-10', 'H5-5', '014Q101046', '014S101194', '014P101430')

### Code that was previously used for determining the above sampled areas
#
# sample_area <- function (area, n) {
#   indices <- sample(nrow(area), n)
#   area[indices,]
# }
#
# vancouver_ramp_sample <- sample_area(vancouver_ramps, 1)
# camas_ramp_sample <- sample_area(camas_ramps, 1)
# both_are_on_ramps <- is_on_ramp(vancouver_ramp_sample) & is_on_ramp(camas_ramp_sample)
# both_are_off_ramps <- is_off_ramp(vancouver_ramp_sample) & is_off_ramp(camas_ramp_sample)
#
# washougal_ramp_sample <- NA
#
# if (both_are_off_ramps) {
#   washougal_ramp_sample <- washgoual_on_ramp
# } else if (both_are_on_ramps) {
#   washougal_ramp_sample <- washougal_off_ramp
# } else {
#   washougal_ramp_sample <- sample_area(washougal_ramps, 1)
# }
#
# samples <- rbind(
#   sample_area(area1N, 1),
#   sample_area(area1S, 1),
#   sample_area(area2S, 1),
#   sample_area(area3N, 1),
#   sample_area(area3S, 1),
#   vancouver_ramp_sample,
#   camas_ramp_sample,
#   washougal_ramp_sample
# )
```

# Testing & Verification

```{r}
verify_segment_lengths <- function (area) {
  segment_lengths <- area |> st_transform(CRS_FEET) |> st_length()
  area_length <- segment_lengths |> sum() / 5280
  c(segment_lengths, area_length)
}

verify_segment_lengths(area1N)
verify_segment_lengths(area1S)
verify_segment_lengths(area2S)
verify_segment_lengths(area3N)
verify_segment_lengths(area3S)

verify_segment_lengths(vancouver_ramps)
verify_segment_lengths(camas_ramps)
verify_segment_lengths(washougal_ramps)
```

# Export Data

```{r}
out_dir = 'out'
if (!dir.exists(out_dir)) {
  dir.create(out_dir)
}

full_data <- data.frame(rbind(area1N, area1S, area2S, area3N, area3S, vancouver_ramps, camas_ramps, washougal_ramps))
data_cols <- colnames(full_data) |> head(-1) # exclude geometry data in the last column
full_data <- full_data[, data_cols]
write.csv(full_data, paste0(out_dir, '/full.csv'))

sample_data <- full_data |> filter(area.id %in% sample_area_ids)
write.csv(sample_data, paste0(out_dir, '/samples.csv'))
```

# Maps

```{r}
map.provider <- 'CartoDB.Positron'
map.credit <- get_credit(map.provider)

get_croppable_bbox <- function (sf_obj, gap) {
  box <- st_bbox(sf_obj |> st_transform(CRS_FEET))
  st_bbox(c(
    box$xmin - gap,
    box$xmax + gap,
    box$ymin - gap,
    box$ymax + gap
  ), crs = st_crs(CRS_FEET))
}

get_map_tiles <- function (sf_obj, zoom=16, crop=F, gap=500.0, forceDownload=F) {
  sf_obj_feet <- sf_obj |> st_transform(CRS_FEET)
  box = st_bbox(sf_obj_feet)
  if (crop) {
    box = get_croppable_bbox(sf_obj_feet, gap)
  }
  tiles <- get_tiles(
    box,
    provider = map.provider,
    zoom = zoom,
    crop = crop,
    forceDownload = forceDownload
  )
}

make_tiles_df <- function (tiles) {
  tiles_df <- as.data.frame(tiles, xy = T)
  colnames(tiles_df) <- c('x', 'y', 'red', 'green', 'blue')
  tiles_df
}

plot_map <- function (tiles, objects, legend, legend_theme) {
  map.df <- make_tiles_df(tiles)
  ggplot() +
    geom_raster(data = map.df, aes(x = x, y = y, fill = rgb(red, green, blue, maxColorValue = 255))) +
    geom_sf(data = objects, aes(color = type, linewidth = line_size)) +
    scale_fill_identity() +
    scale_linewidth_identity() +
    coord_sf() +
    theme_void() +
    annotate(geom = 'text', x = max(map.df$x), y = min(map.df$y), label = map.credit, hjust = 'inward', vjust = 'inward', size = 2) +
    legend +
    legend_theme
}

plot_ramp_map <- function (tiles, objects, legend_position) {
  legend <- scale_color_manual(name = '', values = c('Unsampled' = 'darkgreen', 'Sampled' = 'orange'))
  legend_theme <- theme(legend.position = legend_position, legend.justification = c(1, 1))
  plot_map(tiles, objects, legend, legend_theme)
}

save_map_image <- function (map_data, filename) {
  ggsave(paste0(out_dir, '/', filename, '.png'), plot=map_data, width = 8, height = 8, dpi = 300)
}

clean_sf <- function (sf_data) {
  sf_column <- attr(data, 'sf_column')
  data <- sf_data |>
    st_transform(CRS_FEET) |>
    select({{ sf_column }})
  colnames(data) <- c('geometry')
  attr(data, 'sf_column') <- 'geometry'
  data
}
```

## Highways

```{r}
area1N_full <- extract_milepost_segments(wa14_mp_dec, min(area1N$start.milepost), max(area1N$end.milepost))
area1S_full <- extract_milepost_segments(wa14_mp_inc, min(area1S$start.milepost), max(area1S$end.milepost))
area2S_full <- extract_milepost_segments(wa14_mp_inc, min(area2S$start.milepost), max(area2S$end.milepost))
area3N_full <- extract_milepost_segments(wa14_mp_dec, min(area3N$start.milepost), max(area3N$end.milepost))
area3S_full <- extract_milepost_segments(wa14_mp_inc, min(area3S$start.milepost), max(area3S$end.milepost))
all_highways <- rbind(area1N_full, area1S_full, area2S_full, area3N_full, area3S_full) |>
  st_transform(CRS_FEET)

east_highways <- rbind(area3N_full, area3S_full)
west_highways <- rbind(area1N_full, area1S_full)
extra_highways <- rbind(area2S_full)

highway_samples <- rbind(area1N, area1S, area2S, area3N, area3S) |>
  filter(area.id %in% sample_area_ids) |>
  select(x)
colnames(highway_samples) <- c('geometry')
attr(highway_samples, 'sf_column') <- 'geometry'

west_highway_samples = highway_samples[1:2,]
extra_highway_samples = highway_samples[3:3,]
east_highway_samples = highway_samples[4:5,]

west_highways.tiles <- get_map_tiles(west_highways, zoom=15, crop=T)
east_highways.tiles <- get_map_tiles(east_highways, zoom=15, crop=T)
extra_highway.tiles <- get_map_tiles(extra_highways, zoom=15, crop=T)
```

Generate and save maps:

```{r}
legend_highway <- scale_color_manual(name = '', values = c('Unsampled' = 'blue', 'Sampled' = 'red'))

west_map.data <- rbind(
  clean_sf(area1N_full) |> cbind(type = 'Unsampled', line_size = 0.4),
  clean_sf(area1S_full) |> cbind(type = 'Unsampled', line_size = 0.4),
  clean_sf(west_highway_samples) |> cbind(type = 'Sampled', line_size = 0.5)
)
west_map <- plot_map(west_highways.tiles, west_map.data, legend_highway, theme(legend.position = c(0.2, 0.4), legend.justification = c(1, 1)))
save_map_image(west_map, 'west_highways')

east_map.data <- rbind(
  clean_sf(area3N_full) |> cbind(type = 'Unsampled', line_size = 0.4),
  clean_sf(area3S_full) |> cbind(type = 'Unsampled', line_size = 0.4),
  clean_sf(east_highway_samples) |> cbind(type = 'Sampled', line_size = 0.5)
)
east_map <- plot_map(east_highways.tiles, east_map.data, legend_highway, theme(legend.position = c(0.2, 0.4), legend.justification = c(1, 1)))
save_map_image(east_map, 'east_highways')

extra_map.data <- rbind(
  clean_sf(area2S_full) |> cbind(type = 'Unsampled', line_size = 0.6),
  clean_sf(extra_highway_samples) |> cbind(type = 'Sampled', line_size = 0.75)
)
extra_map <- plot_map(extra_highway.tiles, extra_map.data, legend_highway, theme(legend.position = c(0.2, 0.9), legend.justification = c(1, 1)))
save_map_image(extra_map, 'extra_highway')
```

## Ramps

```{r}
all_ramps = rbind(vancouver_ramps, camas_ramps, washougal_ramps) |>
  st_transform(CRS_FEET)

ramp_samples <- all_ramps |>
  filter(area.id %in% sample_area_ids) |>
  select(x)

all_ramps <- select(all_ramps, x)

vancouver_ramps.tiles <- get_map_tiles(vancouver_ramps, zoom=17, crop=T, gap=150.0)
camas_ramps.tiles <- get_map_tiles(camas_ramps, zoom=17, crop=T, gap=150.0)
washougal_ramp_east.tiles <- get_map_tiles(washougal_ramps[1,], zoom=17, crop=T, gap=150.0)
washougal_ramp_west.tiles <- get_map_tiles(washougal_ramps[2,], zoom=17, crop=T, gap=150.0)
```

Generate and save maps:

```{r}
vancouver_ramps.data <- rbind(
  clean_sf(vancouver_ramps) |> cbind(type = 'Unsampled', line_size = 0.5),
  clean_sf(ramp_samples[1,]) |> cbind(type = 'Sampled', line_size = 0.6)
)
vancouver_ramps.map <- plot_ramp_map(vancouver_ramps.tiles, vancouver_ramps.data, legend_position=c(0.2, 0.3))
save_map_image(vancouver_ramps.map, 'vancouver_ramps')

camas_ramps.data <- rbind(
  clean_sf(camas_ramps) |> cbind(type = 'Unsampled', line_size = 0.5),
  clean_sf(ramp_samples[2,]) |> cbind(type = 'Sampled', line_size = 0.6)
)
camas_ramps.map <- plot_ramp_map(camas_ramps.tiles, camas_ramps.data, legend_position=c(0.2, 1))
save_map_image(camas_ramps.map, 'camas_ramps')

washougal_ramp_east.data <- rbind(
  clean_sf(washougal_ramps[1,]) |> cbind(type = 'Unsampled', line_size = 0.5)
)
washougal_ramp_east.map <- plot_ramp_map(washougal_ramp_east.tiles, washougal_ramp_east.data, legend_position=c(0.2, 0.4))
save_map_image(washougal_ramp_east.map, 'washougal_ramp_east')

washougal_ramp_west.data <- rbind(
  clean_sf(washougal_ramps[2,]) |> cbind(type = 'Unsampled', line_size = 0.5),
  clean_sf(ramp_samples[3,]) |> cbind(type = 'Sampled', line_size = 0.6)
)
washougal_ramp_west.map <- plot_ramp_map(washougal_ramp_west.tiles, washougal_ramp_west.data, legend_position=c(0.2, 0.45))
save_map_image(washougal_ramp_west.map, 'washougal_ramp_west')
```

## All Together

```{r}
colnames(all_ramps) <- colnames(all_highways)
attr(all_ramps, 'sf_column') <- attr(all_highways, 'sf_column')

colnames(ramp_samples) <- colnames(highway_samples)
attr(ramp_samples, 'sf_column') <- attr(highway_samples, 'sf_column')

west_highways <- all_highways[1:2,]
east_highways <- all_highways[3:5,]

west_ramps <- all_ramps[1:4,]
east_ramps <- all_ramps[5:6,]

west_areas <- rbind(west_highways, west_ramps)
east_areas <- rbind(east_highways, east_ramps)

west_areas.tiles <- get_map_tiles(west_areas, zoom=14, crop=T, gap=500)
east_areas.tiles <- get_map_tiles(east_areas, zoom=14, crop=T, gap=500)

unsampled_highway <- 'Unsampled Highway'
sampled_highway <- 'Sampled Highway'
unsampled_ramp <- 'Unsampled Ramp'
sampled_ramp <- 'Sampled Ramp'

legend_full <- scale_color_manual(
  name = '',
  values = c(
    'Unsampled Highway' = 'blue',
    'Sampled Highway' = 'red',
    'Unsampled Ramp' = 'darkgreen',
    'Sampled Ramp' = 'orange'
  ),
  breaks = c(unsampled_highway, sampled_highway, unsampled_ramp, sampled_ramp)
)

legend_theme <- theme(
  legend.position = c(0.93, 0.98),
  legend.justification = c(1, 1),
  legend.text = element_text(size = 8),
  legend.key.spacing.y = unit(-0.3, 'lines')
)
line_size_full = 0.3

west_map_data <- rbind(
  clean_sf(west_highways) |> cbind(type = unsampled_highway, line_size = line_size_full),
  clean_sf(highway_samples[1:2,])|> cbind(type = sampled_highway, line_size = line_size_full),
  clean_sf(west_ramps) |> cbind(type = unsampled_ramp, line_size = line_size_full),
  clean_sf(ramp_samples[1:2,]) |> cbind(type = sampled_ramp, line_size = line_size_full)
)
west_areas.map <- plot_map(west_areas.tiles, west_map_data, legend_full, legend_theme)
save_map_image(west_areas.map, 'full_map_west')

east_map_data <- rbind(
  clean_sf(east_highways) |> cbind(type = unsampled_highway, line_size = line_size_full),
  clean_sf(highway_samples[3:5,]) |> cbind(type = sampled_highway, line_size = line_size_full),
  clean_sf(east_ramps) |> cbind(type = unsampled_ramp, line_size = line_size_full),
  clean_sf(ramp_samples[3,]) |> cbind(type = sampled_ramp, line_size = line_size_full)
)
east_areas.map <- plot_map(east_areas.tiles, east_map_data, legend_full, legend_theme)
save_map_image(east_areas.map, 'full_map_east')

legend_full_no_samples <- scale_color_manual(
  name = '',
  values = c(
    'Highway' = 'blue',
    'Ramp' = 'darkgreen'
  ),
  breaks = c('Highway', 'Ramp')
)

west_areas_no_samples <- west_map_data |>
  filter(type %in% c(unsampled_highway, unsampled_ramp)) |>
  mutate(type = str_remove(type, 'Unsampled '))
west_areas_no_samples.map <- plot_map(west_areas.tiles, west_areas_no_samples, legend_full_no_samples, legend_theme)
save_map_image(west_areas_no_samples.map, 'full_map_west_no_samples')

east_areas_no_samples <- east_map_data |>
  filter(type %in% c(unsampled_highway, unsampled_ramp)) |>
  mutate(type = str_remove(type, 'Unsampled '))
east_areas_no_samples.map <- plot_map(east_areas.tiles, east_areas_no_samples, legend_full_no_samples, legend_theme)
save_map_image(east_areas_no_samples.map, 'full_map_east_no_samples')
```

